A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates

### 2019-07-31 paper.tex
Updates from Overleaf

One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^o$ N, geodetic longitude: 144.189$^o$ E) were obtained from the Japan Meteorological Agency data portal. The location of the observatory and substation are shown in Figure~\ref{map}.

@@

\begin{table}
  \caption{Out-of-sample model performance metrics for Model $m$. MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
  \centering
  \begin{tabular}{l c c c c}
    \hline
    $m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
    \hline
    1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
    2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
    3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
    4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
    \hline
  \end{tabular}
  \label{results}
\end{table}

@@

Note that the values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction. \cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2 degrees west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11 degrees west of north.

@@

The ratio of $|\overline{b}_o|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}. The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.

Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive coefficients and 180$^o$ for negative coefficients by definition). The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different. At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180 degrees and the associated signs are consistent with those of $\overline{a}_o$ and $\overline{b}_o$.

%At $T=12$~h, the phases of $|a(\omega)|$ and $|b(\omega)|$ match that of $|\overline{b}_o|$ and $|\overline{a}_o|$ and the ratio of  $|b(\omega)|/|a(\omega)|$ of 1.6 is the closest it gets to that of $|\overline{b}_o|/|\overline{a}_o|$ of approximately 1. However, the absolute values of $a(\omega)$ and $b(\omega)$ are significantly larger than $a_o$ and $b_o$

%From Figure~\ref{Z} and Figure~\ref{Phi}, it follows that the In Figure~\ref{Phi}, at $T=12$~h, the phase difference between $a(\omega)$ and $b(\omega)$ of 180 degrees is the same as that for $a_o$ and $b_o$. In addition the ratio of $|b(\omega)|/|a(\omega)| =1.6$ is nearest to the frequency independent value predicted by Model~1 of $|\overline{b}_o|/|\overline{a}_o| = 1$.

@@

Note that in this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity. Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line (for both Model~1 and Model~2) is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2. Testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.

%We have argued above that a spatial variation in the spectra of the electric field can explain the frequency dependence of $a$ and $b$. There is also evidence that the electric field varies along

%A consideration of the ratio of $b_o/a_o$ suggests that the ground conductivity varies significantly along the length of the line. As shown in Figure~\ref{map}, the 187~kV line runs in a direction that is approximately 45$^o$ east of north.  Assuming negligible GIC contributions from other unknown parts of the network and an electric field that is uniform along the length of the line, the equation for Model~1, $GIC(t) = a_oE_x(t) + b_oE_y(t)$, where $x$ is northward and $y$ is eastward,

%one would expect the values of $a_o$ and $b_o$ to be such that $tan^{1}b_o/a_o$ is approximately 45$^o$ east of north .

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
  \caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) Geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB). The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, and the date labels are for Universal Time.}
  \label{sample}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-2006-08-06.pdf}
  \caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval. Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
  \label{predictions}
\end{figure}

@@

  %\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
  \caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4. The phase for $a_o$ and $b_o$ are constant and either 0$^o$ or 180$^o$ depending on the sign of their value, with positive values having a phase of 0$^o$.}
  \label{Phi}
\end{figure}

>>>>

One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^{\circ}$ N, geodetic longitude: 144.189$^{\circ}$ E) were obtained from the Japan Meteorological Agency data portal. The location of the observatory and substation are shown in Figure~\ref{map}.

@@

\begin{table}
  \caption{Out-of-sample model performance metrics for Model $m$. MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
  \centering
  \begin{tabular}{l c c c c}
    \hline
    $m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
    \hline
    1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
    2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
    3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
    4\hspace{1em} $G_{B}(t) = z_xB_x(t) + z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
    \hline
  \end{tabular}
  \label{results}
\end{table}

@@

The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) and Model~1 imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction. \cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2$^{\circ}$ west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11$^{\circ}$ west of north.

%The values of $a_o$ and $b_o$ are highly sensitive to the method number of points used in the regression. The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) were found using 86400 points to compute 35 pairs of $a_o$ and $b_o$ values. If 864 points are used to compute 3500 pairs of $a_o$ and $b_o$ values, $\overline{a}_o=+25.7$ and $\overline{b}_o=+15.2$ A/(V/km) and Model~1 predicts maximum GIC when the electric field is approximately 30 degrees east of north, in the approximate direction of the 187 kV line.

@@

The ratio of $|\overline{b}_o|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}. The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h. 
Figure~\ref{Phi} shows the transfer function coefficient phase angles (that for Model~1 is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition). The phases for $a(\omega)$ and $b(\omega)$ are frequency dependent. At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180$^{\circ}$ and the associated signs are consistent with those of $\overline{a}_o > 0$ and $\overline{b}_o < 0$. At $T=12$~h, the values of $b(\omega)$ and $a(\omega)$ and Model~2 correspond to an electric field of 60$^{\circ}$ west of north being associated with the largest GIC.

Both Model~1 and 2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line. This can be caused by part other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity.

@@

In this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity. Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line for both Model~1 and Model~2 is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2. However, testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
  \caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) Geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB). (Bottom) GIC measurements at Memanbetsu substation. The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, the date labels are for Universal Time, and the average have been subtracted from each time series.}
  \label{sample}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-2006-08-06.pdf}
  \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval. Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
  \label{predictions}
\end{figure}

@@

  %\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
  \caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4. The phase for $a_o$ and $b_o$ are constant and either 0$^{\circ}$ or 180$^{\circ}$ depending on the sign of their value, with positive values having a phase of 0$^{\circ}$.}
  \label{Phi}
\end{figure}
